---
title: "Project 2 - 2 of 3"
author: "ZIXIAN LIANG"
date: "2024-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r include=FALSE}
library(tidyverse)
library(readr)
library(tidyr)
library(tibble)
library(dplyr)
library(countrycode)
library(ggplot2)
```


## Dataset 2 - 2023 Euro Long-term interest Rate

### Data Acquisition 


Input file:"Euro Long-term interest.csv” from https://github.com/ZIXIANNOW/DATA607project2/blob/main/Euro%20Long-term%20interest.csv




```{r}
txturl <- "https://raw.githubusercontent.com/ZIXIANNOW/DATA607project2/main/Euro%20Long-term%20interest.csv"
rawdata2 <- read.csv(txturl,  header = TRUE) 
rawdata2
```


### Data Transformation 


Since the dataset solely includes ISO Series keys, it's necessary to extract country abbreviations and create a column with the complete country names for clarity.



```{r}
rawdata2$EU_ISO <- gsub("^(?:[^\\.]*\\.){2}([^\\.]*)\\..*$", "\\1", rawdata2$Series.key)
rawdata2 <- rawdata2[, c("EU_ISO", names(rawdata2)[-which(names(rawdata2) == "EU_ISO")])]
rawdata2$Country <- countrycode(rawdata2$EU_ISO, "iso2c", "country.name")
rawdata2 <- rawdata2[, c("Country", names(rawdata2)[-which(names(rawdata2) == "Country")])]
rawdata2 <- rawdata2 %>% arrange(Country, `Obs..Date`)
rawdata2 <- rawdata2 %>% rename(Month = `Obs..Date`, Interest_Rate = `Obs..value`)
interestdata <- select(rawdata2,Country,EU_ISO,Month,Interest_Rate,Series.key)
interestdata <- separate(interestdata, Month, into = c("Year", "Month"), sep = "-")
interestdata
```



```{r}
ggplot(data = interestdata, aes(x = Month, y = Interest_Rate, color = Country)) +
  geom_point() +
  labs(x = "Month", y = "Interest Rate", title = "2023 Interest Rate by Country and Month")
```


```{r}
filtered_raw <- rawdata2 %>%
  filter(Month %in% c("2023-01", "2023-12"))
filtered_data <- select(filtered_raw,Country,EU_ISO,Month,Interest_Rate,Series.key)
filtered_data

```

```{r}
filtered_update <-  filtered_data %>% 
             pivot_wider(names_from = Month, values_from = Interest_Rate)

filtered_update <- filtered_update %>%
  mutate(Change = `2023-12` - `2023-01`,
         Percentage_Change = Change / `2023-01` * 100) %>%
  arrange(Percentage_Change)

filtered_update
```
```{r}
summary(filtered_update$Percentage_Change)
```


```{r}
ggplot(data = filtered_update, aes(x = Country, y = Percentage_Change)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Country", y = "Percentage Change", title = "2023 Interst Rate Percentage Change by Country") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### Conclusion


In summary, Hungary consistently has the highest interest rate among all EU countries. Over the course of the year, Greece experienced the largest percentage decrease, falling by 23.54%. Lithuania maintained the same interest rate from January to December, making it the most stable in terms of interest rates. Additionally, Bulgaria saw the largest percentage increase, rising by 116.76%.

